\texttt{davos}: a Python package ``smuggler'' for constructing lightweight reproducible notebooks

### 2022-10-26 paper/main.tex
text updates through pg 5 (excluding figs and captions)

% ------------------------------------------------ SOFTWARE DESCRIPTION ------------------------------------------------
\section{Software description}
% Describe the software in as much as is necessary to establish a vocabulary needed to explain its impact.

\begin{figure}[tp]
\centering
\includegraphics[width=\textwidth]{figs/package_structure}
\caption{\small \textbf{Package structure.}}
\label{fig:package-structure}
\end{figure}

\subsection{Software architecture}
% Give a short overview of the overall software architecture; provide a pictorial component overview or similar (if possible). If necessary provide implementation details.
The \texttt{davos} package consists of two interdependent subpackages (see Fig.~\ref{fig:package-structure}). The first, \texttt{davos.core}, comprises a set of modules that implement\comment{provide?} the bulk of the package's core functionality, including pipelines for installing and validating packages, custom parsers for the \texttt{smuggle} statement (see Section~\ref{subsec:smuggle}) and onion comment (see Section~\ref{subsec:onion}), and a runtime interface for configuring \texttt{davos}'s behavior (see Section~\ref{subsec:config}). However, certain critical\comment{other important} aspects of this functionality require (often substantially) different implementation approaches depending on various properties of the notebook environment in which \texttt{davos} is used (e.g., whether the frontend is provided by Jupyter or Google Colaboratory, or which version of IPython~\cite{PereGran07} is used by the notebook kernel). To deal with this, environment-dependent parts of core features and behaviors are isolated and abstracted to ``helper functions'' in the \texttt{davos.implementations} subpackage. This second subpackage defines multiple, interchangeable versions of each helper function, organized into modules by the conditions that trigger their use. At runtime, \texttt{davos} detects various features in the notebook environment and selectively imports a single version of each helper function into the top-level \texttt{davos.implementations} namespace, allowing \texttt{davos.core} modules to access the correct implementations for the current notebook environment in a single, consistent location. An additional benefit of this design pattern is that it allows maintainers or users to easily extend \texttt{davos} to support new, updated, or custom notebook variants simply by creating a new \texttt{davos.implementations} module with any necessary tweaks to existing helper functions.

@@

\subsection{Software functionalities}
% Present the major functionalities of the software.
\subsubsection{The \texttt{smuggle} statement}\label{subsec:smuggle}
Importing \texttt{davos} in an IPython notebook enables an additional Python keyword: ``\texttt{smuggle}'' (see Section~\ref{subsec:implementation} for details on how this works).
The \texttt{smuggle} statement can be used as a drop-in replacement for Python's built-in \texttt{import} statement to load libraries, modules, and other objects into the current namespace.
However, whereas \texttt{import} will fail if the requested package is not installed locally, \texttt{smuggle} statements can handle missing packages on the fly.
If a smuggled package does not exist in the local environment, \texttt{davos} will install it automatically, expose its contents to Python's \texttt{import} machinery, and load it into the namespace for immediate use.

\subsubsection{The onion comment}\label{subsec:onion}
For greater control over the behavior of \texttt{smuggle} statements, \texttt{davos} defines an additional construct called the ``onion comment''. An onion comment is a special type of inline comment that may be placed on a line containing a \texttt{smuggle} statement to customize how \texttt{davos} searches for the smuggled package locally and, if necessary, downloads and installs it. Onion comments follow a simple syntax based on the ``type comment'' syntax introduced in PEP 484~\cite{vanREtal14}, and are designed to make managing packages with \texttt{davos} intuitive and familiar. To construct an onion comment, simply provide the name of the installer program (e.g., \texttt{pip}) and the same arguments one would use to manually install the package as desired via the command line:
\begin{center}
\includegraphics[width=0.9\textwidth]{figs/snippet1}
\end{center}
Onion comments are useful when smuggling a package whose distribution name (i.e., the name used when installing it) is different from its top-level module name (i.e., the name used when importing it):
\begin{center}
\includegraphics[width=0.9\textwidth]{figs/snippet2}
\end{center}
However, the most powerful use of the onion comment is making \texttt{smuggle} statements \textit{version-sensitive}. If an onion comment includes a version specifier~\cite{CoghStuf13}, \texttt{davos} will ensure that the version of the package loaded into the notebook matches the specific version requested, or satisfies the given version constraints. If the smuggled package exists locally, \texttt{davos} will extract its version info from its metadata and compare it to the specifier provided. If the two are incompatible (or no local installation is found), \texttt{davos} will install and load a suitable version of the package instead:
\begin{center}
\includegraphics[width=0.9\textwidth]{figs/snippet3}
\end{center}
Onion comments can similarly be used to smuggle specific VCS references (e.g., Git~\cite{TorvHama05}  branches, commits, tags, etc.):
\begin{center}
\includegraphics[width=0.9\textwidth]{figs/snippet4}
\end{center}
\texttt{davos} processes onion comments internally before forwarding arguments to the installer program. In addition to preventing onion comments from being used as a vehicle for shell injection attacks, this allows \texttt{davos} take certain logical actions when particular arguments are passed. For example, the \texttt{-I}/\texttt{--ignore-installed}, \texttt{-U}/\texttt{--upgrade}, and \texttt{--force-reinstall} flags will all cause \texttt{davos} to skip searching for a smuggled package locally before installing a new copy:
\begin{center}
\includegraphics[width=0.9\textwidth]{figs/snippet5}
\end{center}
Similarly, passing \texttt{--no-input} will temporarily enable \texttt{davos}'s non-interactive mode (see Section~\ref{subsec:onion}), and installing a smuggled package into \texttt{<dir>} with \texttt{--target <dir>} will cause \texttt{dir} to be prepended to the module search path (\texttt{sys.path}), if necessary, so the package can be imported

>>>>

% ------------------------------------------------ SOFTWARE DESCRIPTION ------------------------------------------------
\section{Software description}
% Describe the software in as much as is necessary to establish a vocabulary needed to explain its impact.
The \texttt{davos} package is named after Davos Seaworth, a smuggler often referred to as ``the Onion Knight" from the series \textit{A Song of Ice and Fire} by George R. R. Martin.

\begin{figure}[tp]
\centering
\includegraphics[width=\textwidth]{figs/package_structure}
\caption{\small \textbf{Package structure.} }
\label{fig:package-structure}
\end{figure}

\subsection{Software architecture}
% Give a short overview of the overall software architecture; provide a pictorial component overview or similar (if possible). If necessary provide implementation details.
The \texttt{davos} package consists of two interdependent subpackages (see Fig.~\ref{fig:package-structure}). The first, \texttt{davos.core}, comprises a set of modules that implement\comment{provide?} the bulk of the package's core functionality, including pipelines for installing and validating packages, custom parsers for the \texttt{smuggle} statement (see Section~\ref{subsec:smuggle}) and onion comment (see Section~\ref{subsec:onion}), and a runtime interface for configuring \texttt{davos}'s behavior (see Section~\ref{subsec:config}). However, certain critical\comment{other important} aspects of this functionality require (often substantially) different implementations depending on properties of the notebook environment in which \texttt{davos} is used (e.g., whether the frontend is provided by Jupyter or Google Colaboratory, or which version of IPython~\cite{PereGran07} is used by the notebook kernel). To deal with this, environment-dependent parts of core features and behaviors are isolated and abstracted to ``helper functions'' in the \texttt{davos.implementations} subpackage. This second subpackage defines multiple, interchangeable versions of each helper function, organized into modules by the conditions that trigger their use. At runtime, \texttt{davos} detects various features in the notebook environment and selectively imports a single version of each helper function into the top-level \texttt{davos.implementations} namespace, allowing \texttt{davos.core} modules to access the correct implementations for the current notebook environment in a single, consistent location. An additional benefit of this design pattern is that it allows maintainers or users to easily extend \texttt{davos} to support new, updated, or custom notebook variants by creating a new \texttt{davos.implementations} module with any necessary tweaks to the existing helper functions.

@@

\subsection{Software functionalities}
% Present the major functionalities of the software.
\subsubsection{The \texttt{smuggle} statement}\label{subsec:smuggle}
Importing \texttt{davos} in an IPython notebook appears to enable an additional Python keyword: ``\texttt{smuggle}'' (see Section~\ref{subsec:implementation} for details on how this works).
The \texttt{smuggle} statement can be used as a drop-in replacement for Python's built-in \texttt{import} statement to load libraries, modules, and other objects into the current namespace.
However, whereas \texttt{import} will fail if the requested package is not installed locally, \texttt{smuggle} statements can handle missing packages on the fly.
If a smuggled package does not exist in the local environment, \texttt{davos} will install it automatically, expose its contents to Python's \texttt{import} machinery, and load it into the namespace for immediate use.

\subsubsection{The onion comment}\label{subsec:onion}
For greater control over the behavior of \texttt{smuggle} statements, \texttt{davos} defines an additional construct called the ``onion comment.'' An onion comment is a special type of inline comment that may be placed on a line containing a \texttt{smuggle} statement to customize how \texttt{davos} searches for the smuggled package locally and, if necessary, downloads and installs it. Onion comments follow a simple format based on the ``type comment'' syntax introduced in PEP 484~\cite{vanREtal14}, and are designed to make managing packages with \texttt{davos} intuitive and familiar. To construct an onion comment, users provide the name of the installer program (e.g., \texttt{pip}) and the same arguments one would use to manually install the package as desired via the command line:
\begin{center}
\includegraphics[width=0.9\textwidth]{figs/snippet1}
\end{center}
Onion comments are useful when smuggling a package whose distribution name (i.e., the name used when installing it) is different from its top-level module name (i.e., the name used when importing it):
\begin{center}
\includegraphics[width=0.9\textwidth]{figs/snippet2}
\end{center}
However, the most powerful use of the onion comment is making \texttt{smuggle} statements \textit{version-sensitive}. If an onion comment includes a version specifier~\cite{CoghStuf13}, \texttt{davos} will ensure that the version of the package loaded into the notebook matches the specific version requested, or satisfies the given version constraints. If the smuggled package exists locally, \texttt{davos} will extract its version info from its metadata and compare it to the specifier provided. If the two are incompatible (or no local installation is found), \texttt{davos} will install and load a suitable version of the package instead:
\begin{center}
\includegraphics[width=0.9\textwidth]{figs/snippet3}
\end{center}
Onion comments can similarly be used to smuggle specific VCS references (e.g., Git~\cite{TorvHama05}  branches, commits, tags, etc.):
\begin{center}
\includegraphics[width=0.9\textwidth]{figs/snippet4}
\end{center}
\texttt{davos} processes onion comments internally before forwarding arguments to the installer program. In addition to preventing onion comments from being used as a vehicle for shell injection attacks, this allows \texttt{davos} take certain logical actions when particular arguments are passed. For example, the \texttt{-I}/\texttt{--ignore-installed}, \texttt{-U}/\texttt{--upgrade}, and \texttt{--force-reinstall} flags will all cause \texttt{davos} to skip searching for a smuggled package locally before installing a new copy:
\begin{center}
\includegraphics[width=0.9\textwidth]{figs/snippet5}
\end{center}
Similarly, passing \texttt{--no-input} will temporarily enable \texttt{davos}'s non-interactive mode (see Section~\ref{subsec:onion}), and installing a smuggled package into \texttt{<dir>} with \texttt{--target <dir>} will cause \texttt{dir} to be prepended to the module search path (\texttt{sys.path}), if necessary, so the package can be imported
